---
title: "Lab 5"
author: "Hannah Lo/Owner(halo-ds) and Yuma Anderson(yumaanderson)"
date: "10/31/2019"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
```

## Data Preprocess

We are working with a diabetes dataset that contains 19 variables from 403 observations. 
```{r}
data = read.table('diabetes.txt',header=TRUE)
head(data)
```

Missing values in the variable 'frame' were replaced with 'NA'.
```{r}
data$frame[data$frame==''] = NA
data$frame = droplevels(data$frame)
table(data$frame)
```

The variables 'id','bp.2s', and 'bp.2d' were dropped from the dataframe. 
```{r}
diabetes_reduced = select(data,-'id',-'bp.2s',-'bp.2d')
```

Cleanaed the dataset by removing any rows with NA values. 
```{r}
index.na=apply(is.na(diabetes_reduced), 1, any)
diabetes_clean = diabetes_reduced[-which(index.na),]
```

To check that all NA values were removed, the number of rows that contain NA values were computed. 
```{r}
index.check=apply(is.na(diabetes_clean), 1, any)
which(index.check)
```

## Exploring and Transforming Data

The variable 'glyhb' is right skewed. We can remedy this by using the logarthmic function on the 'glyhb' variable. Some potential downsides to this approach is ________
```{r}
diabetes_clean %>% ggplot(aes(glyhb))+ geom_histogram(binwidth = 1)
```

The variable 'glyhb_star' was created by taking the log and is plotted in the following to show its new distribution: 
```{r}

```

Summary statistics 
```{r}

```

## Visualizations 

